A Typed Programmatic Interface to Contracts on the Blockchain

### 2021-01-28 papers/execution-model/model.tex
first cut on motivation

\title{Blockchain System}

@@

\begin{document}
\maketitle
\pagestyle{plain}

@@

Example pseudo code
\begin{align*}
  \OPH &=  \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR \\
       & \MATCH\ {status\ \OPH}\ \WITH \\
       & \mid \STATUSPENDING \to \textrm{not yet included, continue waiting} \\
       & \mid \STATUSINCLUDING (n) \to \textrm{included with $n$
         blocks confirmation}\\
       & \mid \STATUSTIMEOUT \to \textrm{lost, try again}
\end{align*}
Grammar of types
\begin{align*}
  \TYPE & ::= \TADDR \mid \TOPH \mid \TPUH\ \TYPE\ \TYPE \mid \TPUK \mid \TCODE\
          \TYPE\ \TYPE \mid \TTEZ \mid
          \TINT \mid \TUNIT \mid \TBOOL \mid \TSTRING \mid \TSTATUS \\
  & \mid \TYPE\to\TYPE \mid \TPAIR\ \TYPE\ \TYPE \mid \TLIST\ \TYPE
    \mid \TSUM\ \TYPE\ \TYPE \mid \TOPTION\ \TYPE \mid \TEXCEPTION
\end{align*}

>>>>

\title{A Typed Programmatic Interface to Contracts on the Blockchain}

@@

\begin{document}
\maketitle
\pagestyle{plain}
\begin{abstract}
  Integrating blockchain contracts in software systems requires a
  programmatic interface. We propose a typed interface for this
  purpose and establish some properties of the combined system. A
  prototype of the interface is implemented in OCaml for the Tezos
  blockchain. 
\end{abstract}

\section{Motivation}
\label{sec:motivation}

An auctioneer wants to run an auction on the blockchain using a smart
contract. Once to the contract is deployed on the blockchain, the
auction is open for bidding, but it should be closed after 48 hours,
say. Bidding on the auction and closing the auction are two different
entrypoints / methods of the auction contract.

While it is acceptable for the auctioneer to deploy the contract manually, it
is desirable to have a mechanism that reliably closes the auction
after 48 hours have passed. So the auctioneer wants to write a program
to do so.

On the other side, a bidder may want to implement a certain bidding
strategy that starts at a certain amount and increases the bid until a
certain limit is reached. Of course, the bidder only increases if
someone else placed a higher bid. So the bidder also wants to write a
program to implement this strategy.

Both tasks cannot be implemented as smart contracts because they react
to external triggers. Closing requires
to schedule the activation of a contract at a certain time and bidding
requires to watch the state of the contract (current highest bid) and
react if another bidder places a higher bid.

On another level, running a contract requires to propose a fee to the
blockchain, which may or may not be accepted. In any case, starting a
contract may timeout for a range of different reasons. So just
starting a contract with a fixed fee does not guarantee the contract's execution.
This scenario can also be addressed by a program that tries to run the
contract with successively increasing fees until the blockchain
accepts the contract invocation. The program can also calculate the
required fee by observing the current state of the blockchain.

(From this scenario we see the required functionality for a
programmatic interface to a blockchain:)

\begin{itemize}
\item query certain components like the state of a contract (to obtain
  the current highest bid, to check the fees of the last block)
\item invoke contracts and initiate transfers
\end{itemize}

@@

Example pseudo code
\begin{align*}
  \OPH &=  \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR \\
       & \MATCH\ {status\ \OPH}\ \WITH \\
       & \mid \STATUSPENDING \to \textrm{not yet included, continue waiting} \\
       & \mid \STATUSINCLUDING (n) \to \textrm{included with $n$
         blocks confirmation}\\
       & \mid \STATUSTIMEOUT \to \textrm{lost, try again}
\end{align*}
Grammar of types
\begin{align*}
  \TYPE, \TYPEU & ::= \TADDR \mid \TOPH \mid \TPUH\ \TYPE\ \TYPEU \mid
                  \TPUK \mid
                  \TCODE\ \TYPE\ \TYPEU \mid \TTEZ \mid
          \TINT \mid \TUNIT \mid \TBOOL \mid \TSTRING \mid \TSTATUS \\
  & \mid \TYPE\to\TYPEU \mid \TPAIR\ \TYPE\ \TYPEU \mid \TLIST\ \TYPE
    \mid \TSUM\ \TYPE\ \TYPEU \mid \TOPTION\ \TYPE \mid \TEXCEPTION
\end{align*}
